### 2016-06-23 model.tex
Update on Overleaf.

\documentclass{article}
\usepackage[utf8]{inputenc}
\begin{document}

\title{Circular Economy\\
\textit{Model Formulation}
}

\maketitle

%%%%%%%%%%%%%%%%%%
\section{Introduction}

% model rationale etc

%%%%%%%%%%%%%%%%%%
\section{Model Description}

\end{document}